Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance and associated polyandry

### 2017-03-18 ms_pre_post.Rmd
Final edits through the Discussion

---
title: Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance and associated polyandry
author: ''
date: "December 15, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

If we're interested in evolution and not selection (hence the phrasing), then we really need to put more energy into looking at the assumptions underlying genetic variance, and other things having to do with quantitative genetics. At the moment, we're just assuming that there is pretty much always enough genetic variance on which selection to act. I've assumed we'd be most interested in selection given three reproductive strategy traits, but if we're interested in evolvability of traits, then I think we're ignoring some key things.

@@

<!--- XXX XXX XXX LEFT OFF HERE XXX XXX XXX --->

Evolution of reproductive strategies cannot be presumed to occur in isolation from each other in real-world populations, and is instead likely to involve multiple interacting phenotypes resulting in the potential for feedback and degeneracy (i.e., functional redundance) among phenotypes under selection. For example, strong inbreeding depression in offspring fitness is widely hypothesised to drive selection for inbreeding avoidance, which could consequently feed back to drive selection for polyandry [e.g., @Zeh1997; @Jennions2000; @Tregenza2002]. Evolution of inbreeding avoidance can be driven by selection on both pre-copulatory and post-copulatory mechanisms, so simultaneous evolution of each phenotype is likely to be defined by degeneracy, in addition to any direct costs on phenotypes. We used individual-based modelling to explore evolution of pre-copulatory and post-copulatory inbreeding avoidance and polyandry given (1) hard constraints on evolution of post-copulatory inbreeding avoidance, (2) asymmetric costs of pre-copulatory and post-copulatory inbreeding avoidance, and (3) evolution of one inbreeding avoidance phenotype given the pre-existence the other. We thereby highlight fundamental but theoretically under-developed relationships between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance.

@@

<!--- I have re-ordered the Discussion to reflect the order of the questions, but this sacrifices something by putting the specifics points ahead of more general questions (i.e., those references to Figure 4 were not related to any of the three questions specifically, but were used to help make a more general point derived from a holistic understanding of the model). Hence, I have created a new section for more general points concerning reproductive strategy, also removing the conclusion. --->

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. We found that when post-copulatory inbreeding avoidance could evolve, selection for polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but our model provides support for this verbal hypothesis and therefore has widespread implications for future studies of evolution of polyandry.

<!--- Something doesn't seem quite right here. Our model doesn't provide support for the verbal hypothesis in the same way that real data would. Perhaps it *clarifies* the verbal hypothesis, *formalises?* --->

@@

Evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance occurred in our model, but was affected by the joint evolution of polyandry and the existence of asymmetric costs. One cost free inbreeding strategy precluded another more costly inbreeding strategy from persisting in a focal population by modifying its effect on inbreeding avodance (Figure 3). Consequently, long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given sufficient cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003; @Reid2015]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which we suggest are also relevant costs for polyandry, and include costs of elevated risks associated with (1) predation or (2) disease transmission, and costs incurred through (3) loss of time or (4) depletion of energy. Another type of cost relevant for polyandry and inbreeding strategy is risk of male harm caused by sexual conflict [e.g., @Arnqvist2005a; @Parker2006]. We assumed that males were passive in mating encounters with females, but inbreeding theory predicts that males should benefit from a higher tolerance of inbreeding than females, leading to sexual conflict over inbreeding [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. Here we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs thereby reflect empirical observations in some populations and are therefore biologically realistic, other types of costs have also been observed and could therefore be considered in future models. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Different types of costs might affect selection both directly and indirectly through interactions between reproductive strategies, thereby affecting evolution of entire reproductive systems. Evolution of individual phenotypes therefore needs to be considered in the broader context of potentially interacting and phenotypes and their costs.

<!--- I don't understand the contrast between the two comments, the first of which suggests that the empirical examples of costs cited are not needed here, and the second of which suggests that we need to be more explicit about why we did things the way we did. Hence, I have revised to clarify that we made the modelling choices that we did with respect to costs *because they are observed in empirical systems*. Of course, the costs that we modelled are not the *only* kinds of 'costs' that are observed in empirical systems (hence the Roberts reference), so other models might want to look into different types of costs. I agree that it would not be at all surprising if a reviewer raised a question about how we modelled the costs (or anything else), but this always seems to be true for modelling and I think it kind of misses the big picture about how models like this try to develop theory. We can't model *every* kind of cost that we can possibly think of (nor would we want to), and we're not modelling a specific system where we might be able to justify some costs being more reasonable than others. We also can't (I don't think) logically show with an IBM that the type of cost will be arbitray with respect to the questions of interest (i.e., any cost will do). In the end, we have to make reasonable decisions about how to model costs and work from there. I think that citing empirical studies shows that our assumptions of cost type are reasonable for some systems (though potentially not others). Had we made different assumptions, I suspect we'd get queries from some readers wondering why we didn't do it a different way. I don't mean to keep belabouring this point, but I think that the point of using a model to develop theory isn't to match your assumptions with empirical reality as closely as possible (if we tried, we'd miss a lot of important stuff!). Assumptions need to be clear and useful for addressing the theoretical question(s) of interest -- in other words, the model needs to change *how* people think, not *what* they think. When I review theory papers, this is the criteria I have in mind, not (necessarily) whether or not the assumptions match how things are in nature (they might deliberately not be, for the better!). The only reason I have repeatedly justified the cost assumption is, really, for fear that readers might not entirely understand the nature of model assumptions in theory, not because I think the specifics of the assumptions that we make actually matter in terms of thinking about the three aspects of reproductive strategy we consider. --->

<!---If we're spending this much time justifying an assumption, then we should be able to explain *why* it's an important assumption *with respect to addressing our questions* (there are hundreds of other assumptions we don't spend time on -- number of loci probably isn't going to affect our take-home message). If we can't, and the answer is that 'we don't know if it's important or not', then we're probably not doing theory very well. Actually, I don't think the nature of the costs is going to be important because it's not going to affect any key hypotheses that might be derived from our model -- the effect of a direct negative cost has a very clear interpretatation; changing how a cost is paid might quantitatively affect the end *net* cost as a counter-balance to selection, but so will changing the exact values of the costs we do use. There's no good reason to believe that the way a cost is paid will affect any of the major conclusions (degeneracy is important, polyandry is more likely to evolve given post-copulatory selection, etc.) except by tweaking the overall net cost countering adaptive benefit of inbreeding avoidance (alternatively, if we *can* think of one, it would be a great opportunity for another theory paper!).  --->

@@

General predictions concering inbreeding avoidance and polyandry
-------------------------------------------------------------------

Post-copulatory inbreeding avoidance is useless for females that are not polyandrous, and most effective for highly polyandrous females that can choose amongst multiple mates to ensure that offspring are fertilised favourably. In contrast, pre-copulatory mechanisms are most salient for females that mate only once and therefore have no post-copulatory opportunity to avoiding inbreeding. This theory is borne out in our simulation results, as selection for post-copulatory inbreeding avoidance was negligible in populations where polyandry did not evolve, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). Selection for pre-copulatory inbreeding avoidance was also typically slightly stronger when polyandry did not evolve (e.g., Figure 3A vs 3B; see also Supporting Information p. S1-1). In addition to polyandry increasing selection for post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

<!---  Diluting the idea of degeneracy throughout the manuscript now makes me worry that some readers will miss it entirely. If they're not reading carefully, they'll miss (or simply forget) the buried definition in the fourth paragraph of the Introduction. Because the concept of degeneracy becomes so important later on, they'll need to back track to remember what it actually means. I'm less a fan of 'flagging' things for this reason -- readers don't really *need* to understand degeneracy completely until now, but here they have to go all the way back to the Introduction to get the full scope of its importance. --->

<!--- By removing the sentence specifying how degeneracy works with respect to inbreeding avoidance, we are asking the readers to recall these specifics from the introduction (or go back and find them again). The exact mechanism by which functional overlap occurs is removed from the more vague 'phenotypic reproductive strategies'. I'm afraid this will lose some readers who might be a bit tired by the end of the Discussion.--->

@@

<!--- I don't understand the comment about 'getting at the genetic scales'.  --->

>>>>

---
title: Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance and associated polyandry
author: ''
date: "19 March 2017"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

If we're interested in evolution and not selection (hence the phrasing), and not wanting to assume that the latter strictly implies the former (as of course is sometimes the case in nature), then we really need to put more energy into looking at the assumptions underlying genetic variance, and other things having to do with quantitative genetics. At the moment, we're just assuming that there is pretty much always enough genetic variance on which selection to act. I've assumed we'd be most interested in selection given three reproductive strategy traits, but if we're interested in evolvability of traits, then I think we're ignoring some key things. Hence my previous concerns about the theoretical question really not being entirely clear here.

@@

Evolution of reproductive strategies cannot be presumed to occur in isolation from each other in real-world populations, and is instead likely to involve multiple interacting phenotypes resulting in the potential for feedback and degeneracy (i.e., functional redundance) among phenotypes under selection. For example, strong inbreeding depression in offspring fitness is widely hypothesised to drive selection for inbreeding avoidance, which could consequently drive selection for polyandry [e.g., @Zeh1997; @Jennions2000; @Tregenza2002]. Evolution of inbreeding avoidance can be driven by selection on both pre-copulatory and post-copulatory mechanisms, so simultaneous evolution of each phenotype is likely to be defined by degeneracy, in addition to any direct costs on phenotypes. We used individual-based modelling to explore evolution of pre-copulatory and post-copulatory inbreeding avoidance and polyandry given (1) hard constraints on evolution of post-copulatory inbreeding avoidance, (2) asymmetric costs of pre-copulatory and post-copulatory inbreeding avoidance, and (3) evolution of one inbreeding avoidance phenotype given the pre-existence the other. We thereby highlight fundamental but theoretically under-developed relationships between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance.

@@

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. We found that when post-copulatory inbreeding avoidance could evolve, selection for and resulting evolution of polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but our model clarifies this verbal hypothesis and therefore has widespread implications for future studies of evolution of polyandry.

@@

Evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance occurred in our model, but was affected by the joint evolution of polyandry and the existence of asymmetric costs. One cost free strategy of inbreeding avoidance precluded another more costly strategy from persisting in a focal population by modifying its effect on inbreeding avoidance (Figure 3). Consequently, long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given sufficient cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which are also relevant for polyandry, and include costs of elevated risks of predation or disease transmission, and costs incurred through time or energy expenditure. Another type of cost relevant for polyandry and inbreeding strategy is risk of male harm caused by sexual conflict [e.g., @Arnqvist2005a; @Parker2006]. We assumed that males were passive in mating encounters with females, but inbreeding theory predicts that males should benefit from a higher tolerance of inbreeding than females, leading to sexual conflict over inbreeding [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. Here we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs thereby reflect empirical observations in some populations and are therefore biologically realistic, other types of costs have also been observed and could therefore be considered in future models. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Different types of costs might affect selection both directly and indirectly through interactions between reproductive strategies, thereby affecting evolution of entire reproductive systems. Evolution of individual phenotypes therefore needs to be considered in the broader context of potentially interacting and phenotypes and their costs.

@@

General predictions concerning inbreeding avoidance and polyandry
-------------------------------------------------------------------

Post-copulatory inbreeding avoidance cannot be realised if females are not polyandrous, and is most effective for highly polyandrous females that can choose amongst multiple mates and therefore bias fertilisation accordingly. In contrast, pre-copulatory inbreeding avoidance mechanisms are most critical for females that mate only once and therefore have no post-copulatory opportunity to avoiding inbreeding. This theory is borne out in our simulation results, as selection for, and consequent evolution of, post-copulatory inbreeding avoidance was negligible in populations where polyandry did not evolve, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). Evolution of pre-copulatory inbreeding avoidance was also typically slightly stronger when polyandry did not evolve (e.g., Figure 3A vs 3B; see also Supporting Information p. S1-1). In addition to polyandry causing evolution of post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

@@
